### 2019-07-29 submission/manuscript.tex

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{}
    \pretitle{\vspace{\droptitle}}
  \posttitle{}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}

\begin{document}

\vspace{35mm}

\section{\texorpdfstring{Proton pump inhibitor administration does not
promote \emph{Clostridium difficile} colonization in a murine
model}{Proton pump inhibitor administration does not promote Clostridium difficile colonization in a murine model}}\label{proton-pump-inhibitor-administration-does-not-promote-clostridium-difficile-colonization-in-a-murine-model}

\vspace{35mm}

Sarah Tomkovich\({^1}\), Nicholas A. Lesniak\({^1}\), Yuan Li\({^1}\),
Lucas Bishop\({^1}\), Madison J. Fitzgerald\({^1}\), Patrick D.
Schloss\textsuperscript{1\(\dagger\)}

\vspace{40mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}

\(1\) Department of Microbiology and Immunology, University of Michigan,
Ann Arbor, MI 48109

\newpage

\linenumbers

\subsection{Abstract}\label{abstract}

Proton pump inhibitor (PPI) use has been associated with microbiota
alterations and susceptibility to \emph{Clostridium difficile}
infections (CDIs) in humans. We assessed how PPI treatment alters the
fecal microbiota and whether PPIs promote CDIs in a mouse model. Mice
receiving the PPI treatment were gavaged with 40 mg/kg of omeprazole
during a 7-day pretreatment phase, the day of \emph{C. difficile}
challenge, and the following 9 days. We found that mice treated with
omeprazole were not colonized by \emph{C. difficile}. When omeprazole
treatment was combined with a single clindamycin treatment, one cage of
mice remained resistant to \emph{C. difficile} colonization, while the
other cage was colonized. Treatming mice with clindamycin and no
omeprazole followed by challenge resulted in \emph{C. difficile}
colonization. 16S rRNA gene sequencing analysis revealed that omeprazole
had minimal impact on the structure of the murine microbiota throughout
the 16 days of PPI exposure. These results suggest PPI treatment alone
is not sufficient to disrupt microbiota resistance to \emph{C. difficle}
infection in mice that are normally resistant in the absence of
antibiotic treatment.

\subsection{Importance}\label{importance}

Antibiotics are a major risk factor for \emph{Clostridium difficile}
infections (CDIs), but other factors may also contribute. In
epidemiological studies, PPI use has been associated with CDI incidence
and recurrence. Proton pump inhibitors (PPIs) have also been associated
with alterations in the human intestinal microbiota in observational and
interventional studies. We evaluated the effects of the PPI omeprazole
on the structure of the murine intestinal microbiota and its ability to
disrupt colonization resistance to \emph{C. difficile}. We found the PPI
treatment had minimal impact on the murine fecal microbiota and did not
promote \emph{C. difficile} colonization. Further studies are needed to
determine whether other factors such as the composition of the starting
bacterial community, comorbidities, and use of additional medications
contribute to the association between PPIs and CDIs seen in humans or
whether aspects of murine physiology may limit its utility to test these
types of hypotheses.

\newpage

Antibiotics have a large impact on the intestinal microbiome and are a
primary risk factor for developing \emph{Clostridium difficile}
infections (CDIs) (1). It is less clear whether other human medications
that impact the microbiota also influence \emph{C. difficile}
colonization resistance. Multiple epidemiological studies have suggested
an association between proton pump inhibitor (PPI) use and incidence or
recurrence of CDIs (2--5). There have also been a number of large cohort
studies as well as interventional clinical trials that have demonstrated
that specific alterations in the intestinal microbiome were associated
with PPI use (4, 6). PPI-associated microbiota changes have been
attributed to the ability of PPIs to increase stomach acid pH which may
promote the survival of oral and pathogenic bacteria (4, 6). Human fecal
microbiota changes with PPI use include increases in Enterococcaceae,
Lactobacillaceae, Micrococcaceae, Staphylococcaceae and Streptococcaceae
and decreases in Ruminococcaceae (6, 7). Several of these taxa have been
associated with C. difficile colonization in humans (8)\\
Unfortunately, most of the studies suggesting a link between PPIs and
\emph{C. difficile} were retrospective and did not evaluate the
microbiome (2, 3, 5). Thus, it is unclear whether the gastrointestinal
microbiome changes associated with PPI use explain the association
between PPIs and CDIs. Additionally, epidemiological studies have a
limited capacity to address potential confounders and comorbidities in
patients that were on PPIs and developed CDIs or recurrent CDIs (2, 5).
Here, we evaluated the impact of a daily high dose PPI treatment on the
murine microbiome and susceptibility to \emph{C. difficile} colonization
in relation to clindamycin, an antibiotic that perturbs the microbiome
enough to allow \emph{C. difficile} to colonize but is mild enough that
\emph{C. difficile} is cleared within 10 days (9).

\textbf{Murine fecal microbiomes were minimally affected by PPI
treatment } To test whether PPI treatment alters the microbiome and
promotes susceptibility to CDIs, we gavaged mice with 40 mg/kg of
omeprazole for 7 days before \emph{C. difficile} challenge (Figure 1A).
A principle coordinates analysis (PCoA) of the Bray-Curtis distances
over the initial 7 days of treatment revealed the bacterial communities
of PPI-treated mice remained relatively unchanged relative to animals
that did not recieve the PPI (Figure 1B). We observed no significant
fluctuations in the relative abundance of those taxa previously shown to
respond to PPI treatment throughout the course of the 16-day experiment
(Figure 1C, Figure S1A-B). We also observed no significant fluctuations
in relative abundances at the family and genus level over the beginning,
day of \emph{C. difficile} challenge, and end of the experiment for the
PPI-treated mice (Figure 1D, Figure S1C). Next, the fecal bacterial
communities of PPI-treated mice were compared to a group of mice that
received a single 10 mg/kg dose of clindamycin and a group that received
both PPI and clindamycin. In contrast to the dynamics of mice only
receiving omeprazole, the microbiomes from the 2 groups of mice that
received clindamycin started to shift away from the PPI-treated mice and
the samples from previous days one day after treatment (Figure 1D).
These results demonstrated that the PPI treatment alone had a minimal
impact on the murine fecal bacterial community relative to the impact of
clindamycin treatment.

\textbf{PPI-treatment did not promote susceptibility to \emph{C.
difficile infection} in mice} Next, we examined whether PPI treatment
altered susceptibility to \emph{C. difficile} infection in mice. After
PPI treatment or clindamycin treatment, mice were challenged with
\emph{C. difficile} 630 spores. While \emph{C. difficile} colonized the
clindamycin treated mice, it was unable to colonize the PPI-treated mice
(Figure 2A). Interestingly, only 1 cage of mice that received both the
PPI and clindamycin were colonized, while the mice in other cage wwere
resistant (Figure 2A). The greatest shifts in bacterial communities
occurred in the clindamycin-treated mice (Figure 2B, Figure S2).
Regardless of whether the mice became colonized, all of the mice had
cleared \emph{C. difficile} within 5 days (Figure 2A), suggesting that
omeprazole did not affect the rate of clearance. Our results suggest
that PPI treatment had no effects on bacterial community resistance to
\emph{C. difficile} colonization in mice. Instead most of the
differences between our 3 treatment groups appear to be driven by
clindamycin administration (Figure 2C-D) and included previously
described decreases in \emph{Alistipes}, \emph{Barnesiella},
Porhyromonadaceae, Ruminococcaceae (1). These findings demonstrated that
high dose PPI treatment did not promote susceptibility to \emph{C.
difficile} colonization.

\subsection{Conclusions}\label{conclusions}

We found the PPI omeprazole did not impact the gut microbiota and did
not promote \emph{C. difficile} infection in mice. Our findings that PPI
treatment had minimal impact on the fecal microbiome were comparable to
another PPI mouse study that indicated PPIs had more of an effect on the
small intestinal microbiota compared to the fecal microbiota (10). The
same group demonstrated PPI treatment increased the stomach pH in mice
(10), which may improve survival of bacteria passing through the
stomach. We did not find significant changes for the taxa observed to be
significantly impacted by PPI use in human studies. However, 3 of the
human-associated taxa were absent or at low abundance in our mice.
Additionally, some of the significance of PPI associations in human
interventional trials appears to be driven by a handful of specific taxa
with overall differences on a PCoA plot difficult to distinguish (11).
The PPI-treated mice fecal communities did not change overall and no
differences in the relative abundances of individual families or genera
were observed over time. While our fecal microbiota findings are
comparable to what's been shown in another mouse study, whether
PPI-induced changes in specific bacterial abundances observed in humans
play a role in CDIs remains to be determined.\\
Although there have been a few \emph{C. difficile} mouse model studies
that have demonstrated PPIs have some effect on CDIs with or without
additional antibiotic treatment (13--15), there were some key
methodological differences between these studies and our own. One group
administered 0.5 mg/kg of the PPI lansoprazole daily for 2 weeks to mice
and then challenged with \emph{C. difficile} demonstrated that PPI
treatment alone resulted in detectable \emph{C. difficile} in the stool
1 week after challenge, but also showed there was detectable \emph{C.
difficile} in mice not treated with antibiotics (13, 14). The presence
of \emph{C. difficile} in mice not treated with antibiotics may be
partly attributed to the higher dose of vegetative cells
(10\textsuperscript{8} CFU) used, but nevertheless makes it difficult to
interpret the extent to which PPI treatment contributed to \emph{C.
difficile} colonization in their mouse model (13, 14). In contrast, we
have previously shown mice from our colony that were not given
antibiotics were resistant to \emph{C. difficile} 630 when challenged
with 10\textsuperscript{3} spores (16) and previous work by our
collaborators demonstrated mice not given antibiotics were resistant to
10\textsuperscript{5} vegetative cells of \emph{C. difficile } VPI 10463
(17). The other PPI CDI mouse study also used a higher dose of
10\textsuperscript{6} spores or 10\textsuperscript{7} vegetative cells
to challenge antibiotic-treated mice or mice treated with both
antibiotics and the PPI esomeprazole (40mg/kg dose) for 2 days and
demonstrated the antibiotic/PPI-treated mice developed more severe CDIs
(15). We tested the same high 40 mg/kg PPI dose and expanded
pre-treatment to 7 days before challenge to test the impact of PPI
treatment alone on our CDI mouse model.\\
Our study extended the previous work examining PPIs and \emph{C.
difficile} in mice by evaluating the contribution of the intestinal
microbiota. We found the PPI omeprazole had no significant impact on
bacterial taxa within the murine intestinal microbiota over the 16-day
experiment and in contrast to previous work, PPIs did not alter \emph{C.
difficile} colonization resistance. 16S rRNA sequencing suggested that
\emph{Streptococcus} and \emph{Enterococcus} are rare genera in our
C57BL/6 mouse colony and variation in these 2 genera have been observed
across other facilities and vendors (18, 19). These genera could be
important contributors to the associations between PPIs and CDIs in
humans, and could be a contributing factor to our observation that PPI
treatment had no effect on \emph{C. difficile} colonization in our CDI
mouse model.\\
Beyond microbiome differences, factors such as age, body mass index,
comorbidities, and use of other medications in human studies may also be
contributing to the association between PPIs and CDI incidence or
recurrence. This study addressed the impact of PPIs with or without
antibiotics on a murine model of CDI, and found PPIs did not promote
\emph{C. difficile} colonization. The epidemiological evidence linking
PPIs to CDIs is primarily from observational studies, which makes
determining causality and whether other risk factors play a role
challenging (20). Future studies are needed to determine whether age,
other comorbidities and bacterial strains that are less common in mice
can increase the risk of CDIs or recurrent CDIs when combined with PPI
treatment.

\subsection{Acknowledgements}\label{acknowledgements}

This research was supported by NIH grant U01AI12455. We would also like
to thank the Unit for Laboratory Animal Medicine at the University of
Michigan for maintaining our mouse colony and providing the
infrastructure and support for performing our mouse expeirments. The
authors are also thankful to members of the Schloss lab for helpful
discussions throughout the process of designing the experiment,
analyzing the results, crafting the figures, and drafting of the
manuscript.

\newpage

\subsection{Materials and Methods}\label{materials-and-methods}

\textbf{Animals} All mouse experiments were performed with 7- to
12-week-old C57BL/6 male and female mice. Each experimental group of
mice was spit between 2 cages with 2-3 mice housed per cage and male and
female mice housed separately. All animal experiments were approved by
the University of Michigan Animal Care and Use Committee (IACUC) under
protocol number PRO00006983.

\textbf{Drug treatments} Omeprazole (Sigma Aldrich) was prepared in a
vehicle solution of 40\% polyethylene glycol 400 (Sigma-Aldrich) in
phosphate buffered saline. Omeprazole was prepared from 20 mg/mL frozen
aliquots and diluted to an 8mg/mL prior to gavage. All mice received 40
mg/kg omeprazole or vehicle solution once per day through the duration
of the experiment with treatment starting 7 days before \emph{C.
difficile} challenge (Figure 1A). One day prior to \emph{C. difficile}
challenge, 2 groups of mice received an intraperitoneal injection of 10
mg/kg clindamycin or sterile saline vehicle. All drugs were filter
sterilized through a 0.22 micron syringe filter before administration to
animals.

\textbf{\emph{C. difficile} infection model} Mice were challenged with
\emph{C. difficile} 630 seven days after the start of omeprazole
treatment and one day after clindamycin treatment. Mice were challenged
with 10\textsuperscript{3} spores in ultrapure distilled water. Stool
samples were collected for 16S rRNA sequencing or \emph{C. difficile}
CFU quantification throughout the duration of the experiments at the
indicated timepoints (Figure 1A). Samples for 16S rRNA sequencing were
flash frozen in liquid nitrogen and stored at -80°C until DNA
extraction, while samples for CFU quantification were transferred into
an anaerobic chamber and serially diluted in PBS. Diluted samples were
plated on TCCFA (taurocholate, cycloserine, cefoxitin, fructose agar)
plates and incubated at 37°C for 24 hours under anaerobic conditions to
quantify \emph{C. difficile} CFU.

\textbf{16S rRNA gene sequencing} DNA for 16S rRNA gene sequencing was
extracted from 10-50 mg fecal pellet from each mouse using the DNeasy
Powersoil HTP 96 Kit (Qiagen) and an EpMotion 5075 automated pipetting
system (Eppendorf). The ZymoBIOMICS\textsuperscript{TM} Microbial
Community DNA Standard (Zymo, CA, USA) was used as a mock community (21)
and water was used as a negative control. The V4 hypervariable region of
the 16S rRNA gene was amplified with Accuprime Pfx DNA polymerase
(Thermo Fisher Scientific) using previously described custom barcoded
primers (22). The 16S rRNA amplicon library was sequenced with the MiSeq
(Illumina). Amplicons were cleaned up and normalized with the SequalPrep
Normalization Plate Kit (ThermoFisher Scientific) and pooled amplicons
were quantified with the KAPA library quantification kit (KAPA
Biosystems).

\textbf{16S rRNA gene sequence analysis}Mothur (v1.40.5) was used for
all sequence processing steps (23) using the previously published
protocol (22). In brief, forward and reverse reads for each sample were
combined and low-quality sequences and chimeras were removed. Duplicate
sequences were merged, before taxonomy assignment using a modified
version (v16) of the Ribosomal Database Project reference database
(v11.5) with an 80\% cutoff. Operational taxonomic units (OTUs) were
assigned with the opticlust clustering algorithm using a 97\% similarity
threshold. To adjust for uneven sequencing across samples, all samples
were rarefied to 3000 sequences, 1,000 times. PCoAs were generated based
on Bray-Curtis distance. R (v.3.5.1) was used to generate figures and
perform statistical analysis.

\textbf{Statistical Analysis} To test for differences in relative
abundances in families and genera across our 3 different treatment
groups (Clindamycin, Clindamycin + PPI, and PPI) or within the PPI
treatment group across 3 timepoints (Day -7, 0, and 9), we used a
Kruskal-Wallis test with a Benjamini-Hochberg correction for multiple
comparisons.

\textbf{Code availability} The code for all sequence processing and
analysis step as well as an Rmarkdown version of this manuscript is
available at
\url{https://github.com/SchlossLab/Tomkovich_PPI_XXXX_2019}.

\textbf{Data availability} The 16S rRNA sequencing data have been
deposited in the NCBI Sequence Read Archive (Accession no. PRJNA554866).

\newpage

\subsection{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-Schubert2015}{}
1. \textbf{Schubert AM}, \textbf{Sinani H}, \textbf{Schloss PD}. 2015.
Antibiotic-induced alterations of the murine gut microbiota and
subsequent effects on colonization resistance against clostridium
difficile. mBio \textbf{6}.
doi:\href{https://doi.org/10.1128/mbio.00974-15}{10.1128/mbio.00974-15}.

\hypertarget{ref-tariq2017association}{}
2. \textbf{Tariq R}, \textbf{Singh S}, \textbf{Gupta A}, \textbf{Pardi
DS}, \textbf{Khanna S}. 2017. Association of gastric acid suppression
with recurrent clostridium difficile infection: A systematic review and
meta-analysis. JAMA internal medicine \textbf{177}:784--791.

\hypertarget{ref-nehra2018proton}{}
3. \textbf{Nehra AK}, \textbf{Alexander JA}, \textbf{Loftus CG},
\textbf{Nehra V}. 2018. Proton pump inhibitors: Review of emerging
concerns, pp. 240--246. \emph{In} Mayo clinic proceedings. Elsevier.

\hypertarget{ref-Naito2018}{}
4. \textbf{Naito Y}, \textbf{Kashiwagi K}, \textbf{Takagi T},
\textbf{Andoh A}, \textbf{Inoue R}. 2018. Intestinal dysbiosis secondary
to proton-pump inhibitor use. Digestion \textbf{97}:195--204.
doi:\href{https://doi.org/10.1159/000481813}{10.1159/000481813}.

\hypertarget{ref-Elias2019}{}
5. \textbf{Elias E}, \textbf{Targownik LE}. 2019. The clinician's guide
to proton pump inhibitor related adverse events. Drugs
\textbf{79}:715--731.
doi:\href{https://doi.org/10.1007/s40265-019-01110-3}{10.1007/s40265-019-01110-3}.

\hypertarget{ref-Imhann2017}{}
6. \textbf{Imhann F}, \textbf{Vila AV}, \textbf{Bonder MJ},
\textbf{Manosalva AGL}, \textbf{Koonen DP}, \textbf{Fu J},
\textbf{Wijmenga C}, \textbf{Zhernakova A}, \textbf{Weersma RK}. 2017.
The influence of proton pump inhibitors and other commonly used
medication on the gut microbiota. Gut Microbes \textbf{8}:351--358.
doi:\href{https://doi.org/10.1080/19490976.2017.1284732}{10.1080/19490976.2017.1284732}.

\hypertarget{ref-maier2018extensive}{}
7. \textbf{Maier L}, \textbf{Pruteanu M}, \textbf{Kuhn M},
\textbf{Zeller G}, \textbf{Telzerow A}, \textbf{Anderson EE},
\textbf{Brochado AR}, \textbf{Fernandez KC}, \textbf{Dose H},
\textbf{Mori H}, \textbf{others}. 2018. Extensive impact of
non-antibiotic drugs on human gut bacteria. Nature \textbf{555}:623.

\hypertarget{ref-Schuberte01021-14}{}
8. \textbf{Schubert AM}, \textbf{Rogers MAM}, \textbf{Ring C},
\textbf{Mogle J}, \textbf{Petrosino JP}, \textbf{Young VB},
\textbf{Aronoff DM}, \textbf{Schloss PD}. 2014. Microbiome data
distinguish patients with clostridium difficile infection and non-c.
difficile-associated diarrhea from healthy controls. mBio \textbf{5}.
doi:\href{https://doi.org/10.1128/mBio.01021-14}{10.1128/mBio.01021-14}.

\hypertarget{ref-Jenior2018}{}
9. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2018. Clostridium difficile alters the structure
and metabolism of distinct cecal microbiomes during initial infection to
promote sustained colonization. mSphere \textbf{3}.
doi:\href{https://doi.org/10.1128/msphere.00261-18}{10.1128/msphere.00261-18}.

\hypertarget{ref-Yasutomi2018}{}
10. \textbf{Yasutomi E}, \textbf{Hoshi N}, \textbf{Adachi S},
\textbf{Otsuka T}, \textbf{Kong L}, \textbf{Ku Y}, \textbf{Yamairi H},
\textbf{Inoue J}, \textbf{Ishida T}, \textbf{Watanabe D}, \textbf{Ooi
M}, \textbf{Yoshida M}, \textbf{Tsukimi T}, \textbf{Fukuda S},
\textbf{Azuma T}. 2018. Proton pump inhibitors increase the
susceptibility of mice to oral infection with enteropathogenic bacteria.
Digestive Diseases and Sciences \textbf{63}:881--889.
doi:\href{https://doi.org/10.1007/s10620-017-4905-3}{10.1007/s10620-017-4905-3}.

\hypertarget{ref-Freedberg2015}{}
11. \textbf{Freedberg DE}, \textbf{Toussaint NC}, \textbf{Chen SP},
\textbf{Ratner AJ}, \textbf{Whittier S}, \textbf{Wang TC}, \textbf{Wang
HH}, \textbf{Abrams JA}. 2015. Proton pump inhibitors alter specific
taxa in the human gastrointestinal microbiome: A crossover trial.
Gastroenterology \textbf{149}:883--885.e9.
doi:\href{https://doi.org/10.1053/j.gastro.2015.06.043}{10.1053/j.gastro.2015.06.043}.

\hypertarget{ref-Imhann2015}{}
12. \textbf{Imhann F}, \textbf{Bonder MJ}, \textbf{Vila AV}, \textbf{Fu
J}, \textbf{Mujagic Z}, \textbf{Vork L}, \textbf{Tigchelaar EF},
\textbf{Jankipersadsing SA}, \textbf{Cenit MC}, \textbf{Harmsen HJM},
\textbf{Dijkstra G}, \textbf{Franke L}, \textbf{Xavier RJ},
\textbf{Jonkers D}, \textbf{Wijmenga C}, \textbf{Weersma RK},
\textbf{Zhernakova A}. 2015. Proton pump inhibitors affect the gut
microbiome. Gut \textbf{65}:740--748.
doi:\href{https://doi.org/10.1136/gutjnl-2015-310376}{10.1136/gutjnl-2015-310376}.

\hypertarget{ref-Kaur2007}{}
13. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2007. Comparative role of antibiotic
and proton pump inhibitor in ExperimentalClostridium difficileInfection
in mice. Microbiology and Immunology \textbf{51}:1209--1214.
doi:\href{https://doi.org/10.1111/j.1348-0421.2007.tb04016.x}{10.1111/j.1348-0421.2007.tb04016.x}.

\hypertarget{ref-kaur2011effect}{}
14. \textbf{Kaur S}, \textbf{Vaishnavi C}, \textbf{Prasad KK},
\textbf{Ray P}, \textbf{Kochhar R}. 2011. Effect of lactobacillus
acidophilus \& epidermal growth factor on experimentally induced
clostridium difficile infection. The Indian journal of medical research
\textbf{133}:434.

\hypertarget{ref-hung2015proton}{}
15. \textbf{Hung Y-P}, \textbf{Ko W-C}, \textbf{Chou P-H}, \textbf{Chen
Y-H}, \textbf{Lin H-J}, \textbf{Liu Y-H}, \textbf{Tsai H-W}, \textbf{Lee
J-C}, \textbf{Tsai P-J}. 2015. Proton-pump inhibitor exposure aggravates
clostridium difficile--associated colitis: Evidence from a mouse model.
The Journal of infectious diseases \textbf{212}:654--663.

\hypertarget{ref-Jenior2017}{}
16. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2017. Clostridium difficile colonizes alternative
nutrient niches during infection across distinct murine gut microbiomes.
mSystems \textbf{2}.
doi:\href{https://doi.org/10.1128/msystems.00063-17}{10.1128/msystems.00063-17}.

\hypertarget{ref-Reeves2011}{}
17. \textbf{Reeves AE}, \textbf{Theriot CM}, \textbf{Bergin IL},
\textbf{Huffnagle GB}, \textbf{Schloss PD}, \textbf{Young VB}. 2011. The
interplay between microbiome dynamics and pathogen dynamics in a murine
model ofClostridium difficileInfection. Gut Microbes
\textbf{2}:145--158.
doi:\href{https://doi.org/10.4161/gmic.2.3.16333}{10.4161/gmic.2.3.16333}.

\hypertarget{ref-Rausch2016}{}
18. \textbf{Rausch P}, \textbf{Basic M}, \textbf{Batra A},
\textbf{Bischoff SC}, \textbf{Blaut M}, \textbf{Clavel T},
\textbf{Gläsner J}, \textbf{Gopalakrishnan S}, \textbf{Grassl GA},
\textbf{Günther C}, \textbf{Haller D}, \textbf{Hirose M},
\textbf{Ibrahim S}, \textbf{Loh G}, \textbf{Mattner J}, \textbf{Nagel
S}, \textbf{Pabst O}, \textbf{Schmidt F}, \textbf{Siegmund B},
\textbf{Strowig T}, \textbf{Volynets V}, \textbf{Wirtz S},
\textbf{Zeissig S}, \textbf{Zeissig Y}, \textbf{Bleich A},
\textbf{Baines JF}. 2016. Analysis of factors contributing to variation
in the c57bl/6J fecal microbiota across german animal facilities.
International Journal of Medical Microbiology \textbf{306}:343--355.
doi:\href{https://doi.org/10.1016/j.ijmm.2016.03.004}{10.1016/j.ijmm.2016.03.004}.

\hypertarget{ref-Robertson2019}{}
19. \textbf{Robertson SJ}, \textbf{Lemire P}, \textbf{Maughan H},
\textbf{Goethel A}, \textbf{Turpin W}, \textbf{Bedrani L},
\textbf{Guttman DS}, \textbf{Croitoru K}, \textbf{Girardin SE},
\textbf{Philpott DJ}. 2019. Comparison of co-housing and littermate
methods for microbiota standardization in mouse models. Cell Reports
\textbf{27}:1910--1919.e2.
doi:\href{https://doi.org/10.1016/j.celrep.2019.04.023}{10.1016/j.celrep.2019.04.023}.

\hypertarget{ref-Eze2017}{}
20. \textbf{Eze P}, \textbf{Balsells E}, \textbf{Kyaw MH}, \textbf{Nair
H}. 2017. Risk factors for clostridium difficile infections an overview
of the evidence base and challenges in data synthesis. Journal of Global
Health \textbf{7}.
doi:\href{https://doi.org/10.7189/jogh.07.010417}{10.7189/jogh.07.010417}.

\hypertarget{ref-Sze2019}{}
21. \textbf{Sze MA}, \textbf{Schloss PD}. 2019. The impact of DNA
polymerase and number of rounds of amplification in PCR on 16S rRNA gene
sequence data. mSphere \textbf{4}.
doi:\href{https://doi.org/10.1128/msphere.00163-19}{10.1128/msphere.00163-19}.

\hypertarget{ref-Kozich2013}{}
22. \textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter NT},
\textbf{Highlander SK}, \textbf{Schloss PD}. 2013. Development of a
dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq illumina sequencing platform.
Applied and Environmental Microbiology \textbf{79}:5112--5120.
doi:\href{https://doi.org/10.1128/aem.01043-13}{10.1128/aem.01043-13}.

\hypertarget{ref-Schloss2009}{}
23. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

\newpage

\subsection{Figures}\label{figures}

\includegraphics{figure_1.pdf} \textbf{Figure 1. PPI treatment had
minimal impact on the murine fecal microbiota.} A. Mouse experiment
timeline and logistics. The PPI omeprazole was administered throughout
the duration of the experiment. Clindamycin was administered 1 day
before \emph{C. difficile} challenge on Day 0. Stools for 16S rRNA
sequencing analysis were collected on the days that are labeled (Day -7,
-5, -3, -1, 0, 1, 2, 3, 4, 5, 7, 9). \emph{C. difficile} CFU in the
stool was quantified daily through 6 days post-infection by anaerobic
culture. B. Principal Coordinates Analaysis (PCoA) of Bray-curtis
distances from stool samples for all treatment groups during the intial
7 days of the experiment. Each symbol represents a stool sample from an
individual mouse on a given day. Transparency of the symbol corresponds
to treatment day. Circles represent samples before antibiotic treatment
(administered on Day -1), while open diamonds represent samples from the
day after antibiotic treatment (Day 0). Mice treated with the PPI
omeprazole or vehicle cluster together and do not change much over time
while mice that receive clindamycin have samples that start to separate
from the rest of the samples the day after antibiotic treatment. C.
Relative abundances of families previously associated with PPI use in
humans do not change much over the 16-day course of treatment with PPIs
(7 days before \emph{C. difficile} challenge through 9 days post
\emph{C. difficile} challenge). D. Boxplots showing the relative
abundances of 7 familes within PPI-treated mice that had the lowest
adjusted P values after comparing across 3 timepoints. There were no
significant differences acoss time for any of the identified families
(All P \textgreater{} 0.05) after analysis by Kruskal-Wallis test with a
Benjamini-Hochberg correction for multiple comparisons. See Fig. S1C for
genera with lowest adjusted P values (All P \textgreater{} 0.05). For
C-D, the grey line indicates the limit of detection.

\newpage

\includegraphics{figure_2.pdf} \textbf{Figure 2. PPI treatment alone
does not promote CDIs in mice.} A. \emph{C. difficile} CFUs/g stool
measured each day post \emph{C. difficile} challenge for clindamycin,
clindamycin/PPI, and PPI-treated mice. Lines represent the mean CFU/g
for each treatment group while points represent CFU/g for individual
mice within each group. The black dashed line indicates the limit of
detection. B. PCoA of of Bray-curtis distances from stool samples
collected after antibiotic treatment (last 9 days of the experiment).
Transparency of the symbol corresponds to treatment day. Symbols
represent the \emph{C. difficile} colonization status of the mice
measured 2 days post-infection. Circles represent resistant mice
(\emph{C. difficile} was undetectable in stool samples), while X-shapes
represent mice that were colonized with \emph{C. difficile}, although
all mice cleared \emph{C. difficile} within 5 days of infection.
PPI-treated fecal samples primarily cluster together throughout the
experiment. C. Genera significantly associated with treatment groups of
stool samples across all sequenced timepoints. D. Families significantly
associated with treatment groups of stool samples across all sequenced
timepoints. C,D data were analyzed by Kruskal-Wallis test with a
Benjamini-Hochberg correction for multiple comparisons. For C-D, the
grey line indicates the limit of detection.

\newpage

\includegraphics{figure_s1.pdf} \textbf{Figure S1. Taxa within PPI
treated mice fluctuate over time with no overall trend in either
direction.} Relative abundance over time for Lactobacillaceae (A) and
Ruminococcaceae (B), 2 of the PPI-associated familes from human PPI
studies across all 3 treatment groups. Each point represents the
relative abundance for an individual mouse, while the lines represent
the mean relative abundances for each treatment group of mice. C.
Boxplots showing the relative abundances of 13 genera within PPI-treated
mice that had the lowest adjusted P values after comparing across 3
timepoints. There were no significant differences acoss time for any of
the identified families (All P = 0.568) after analysis by Kruskal-Wallis
test with a Benjamini-Hochberg correction for multiple comparisons. For
A-C, the grey lines indicate the limit of detection.

\newpage

\includegraphics{figure_s2.pdf} \textbf{Figure S2. Microbiota diversity
and richness decrease with antibiotic treatment but remain relatively
constant with PPI treatment.} A. Boxplots of the Shannon Index (A) and
richness (B) values for each group of mice over 3 timepoints (Day -7, 0,
and 9). Each circle represents the value for an individual mouse.

\end{document}